Paper 1: Instrument development

### 2024-05-28 Rmd/Paper1.Rmd
Improved structure of document

---
title: "Instrument development"
author: "Niklas Karlsen, e-post: ..."
date: "2024-05-28"
bibliography: "../data/meta_data/references.bib"
csl: "../scripts/university-of-gothenburg-apa-7th-edition-swedish-legislations.csl"
output: 
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true

@@

The main purpose of this Rmd-script is to reproduce the results in paper 1. The whole process from reducing the raw data to doing the EFA and CFA are discussed.

This R-markdown describes the analyses used to support the validity and reliability of the inferences based on the measure developed in the thesis. This includes exploratory factor analysis (EFA) and the confirmatory factor analysis (CFA) that was done to get the results in paper 1.

@@

The first stage is described in Paper 1. This computational essay [@odden_using_2023] describes the analysis behind the results in paper 1, and emphasises the quantitative analyses done in the final two stages. This concerns item reduction and extraction of factors (EFA) in stage two, and test of dimensionality (CFA) and tests of reliability and validity in stage three.

@@

@@

## Step 4: Data gathering

!NB Anonymisere raw data (blant annet studiested) før jeg legger det ut offentlig.

@@

Reducing the raw data ...

@@

Factor analysis is based on linear analysis (i.e., ordinary least squares-regresion (OLS)). Assumptions underlying OLS-regresion are according to @christophersen_introduksjon_2018 that the residuals have both an average value of 0, constant variance (i.e,. homoscedasticity) and independence (i.e., no autocorrelation). In addition, the dependent variables should depend linearly on the independent variables, which should have no multicollinearity (i.e., meaning that no residuals(?) of independent variable should be strongly correlated with any of the other independent variables).

In addition, according to @flora_old_2012, it is important to decide whether to treat the data as ordinal or interval, as this has consequences for how the items are screened. Treating the data as interval, requires checking of missing data, linearity (checking for outliers and colinearity, i.e., very strong correlation between variables), and normally distributed residuals, checks which are not necessary to do if the data are treated as ordinal.

@@

When we have as few as five answer alternatives (i.e., 5-pt Likert scale), there are weighty arguments for considering the data as ordinal [@flora_old_2012; @izquierdo_exploratory_2014; @watkins_exploratory_2018]. Five answer alternatives do however begin to imitate an interval scale (i.e., a continous scale). Thus, researchers have also argued for considering 5-pt scales as being continous [@christophersen_introduksjon_2018; @norman_likert_2010]. In order to use the full arsenal of parametric methods (i.e., the properties of the sample are decided by a few parameters), such as maximum likelihood (ML), we decided to consider the data as interval. To support this choice further, we also presented the five alternatives as points on a linear scale in our questionnaire to help the respondents visualise the alternatives as being on an interval scale.

@@

>>>>

---
title: "Paper 1: Instrument development"
author: "Niklas Karlsen, OsloMet, Norway, e-mail: niklaska@oslomet.no"
date: "2024-05-28"
bibliography: "../data/meta_data/references.bib"
csl: "../scripts/university-of-gothenburg-apa-7th-edition-swedish-legislations.csl" # (source: https://bookdown.org/yihui/rmarkdown-cookbook/bibliography.html)
output: 
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
    number_sections: true

@@

The main purpose of this Rmd-script is to reproduce the results in paper 1. The whole process from reducing the raw data to doing the EFA and CFA is discussed. The document is written as a computational essay [@odden_using_2023].

This R-markdown describes the analyses used to support the validity and reliability of the inferences about PSTs' TPACK based on the measure developed in the thesis. This includes exploratory factor analysis (EFA) and the confirmatory factor analysis (CFA) that was done to get the results in paper 1.

@@

The first stage is described in Paper 1. This computational essay describes the analysis behind the results in paper 1, and emphasises the quantitative analyses done in the final two stages. This concerns item reduction and extraction of factors (EFA) in stage two, and test of dimensionality (CFA) and tests of reliability and validity in stage three.

@@

## Imports

```{r}
# libraries
```

@@

## Step 4: Data gathering / cleaning

@@

Reducing the raw data ... (transforming from ods to Rdata)

* Step 4a: convert from ods to Rdata; rename items using TPACK typology; treat missing

@@

Factor analysis is based on linear analysis (i.e., ordinary least squares-regression (OLS)). Assumptions underlying OLS-regression are according to @christophersen_introduksjon_2018 that the residuals have both an average value of 0, constant variance (i.e,. homoscedasticity) and independence (i.e., no autocorrelation). In addition, the dependent variables should depend linearly on the independent variables, which should have no multicollinearity (i.e., meaning that no residuals(?) of independent variable should be strongly correlated with any of the other independent variables). 
In addition, according to @flora_old_2012, it is important to decide whether to treat the data as ordinal or interval, as this has consequences for how the items are screened. Treating the data as interval, requires checking of missing data, linearity (checking for outliers and collinearity, i.e., very strong correlation between variables), and normally distributed residuals, checks which are not necessary to do if the data are treated as ordinal.

* Step 5a: check residuals (normality, autocorrelation)
* Step 5b: check correlations (collinearity, multicollinearity)
* Step 5c: check outliers (skew, kurtosis, range)
* Step 5d: check missing (pattern: MCAR, MAR, NMAR)
* Step 5e: check eigenvalues, item-test correlation, reliability [@izquierdo_exploratory_2014]

@@

When we have as few as five answer alternatives (i.e., 5-pt Likert scale), there are weighty arguments for considering the data as ordinal [@flora_old_2012; @izquierdo_exploratory_2014; @watkins_exploratory_2018]. Five answer alternatives do however begin to imitate an interval scale (i.e., a continous scale). Thus, researchers have also argued for considering 5-pt scales as being continuous [@christophersen_introduksjon_2018; @norman_likert_2010]. In order to use the full arsenal of parametric methods (i.e., the properties of the sample are decided by a few parameters), such as maximum likelihood (ML), we decided to consider the data as interval. To support this choice further, we also presented the five alternatives as points on a linear scale in our questionnaire to help the respondents visualise the alternatives as being on an interval scale.

@@

## Step 6: Factor extraction (i.e., EFA)

# Stage 3

## Step 7: Tests of dimensionality (i.e., CFA)

## Step 8: Tests of reliability

## Step 9: Tests of validity

# Discussion and conclusion

<div id="refs"></div>

<!-- source: https://bookdown.org/yihui/rmarkdown-cookbook/bibliography.html

# (APPENDIX) Appendix {-} -->
